### 2012-12-26 relazioni/ipermediali/template.tex

\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage[utf8x]{inputenc}
\usepackage[italian]{babel}
\usepackage[colorlinks=true,linkcolor=black]{hyperref}
\usepackage{hhline}
\usepackage{hyperref}
\usepackage[font=small,format=plain,labelfont=bf,up,textfont=it,up]{caption}
\usepackage{indentfirst}
\usepackage[Lenny]{fncychap}
\usepackage{listings}
\usepackage{verbatim}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}
\usepackage{booktabs}
\usepackage{tikz}
\usepackage[hang,centerlast]{subfigure}

\newcommand{\makeTitlePage} {
\begin{titlepage}
\begin{center}
\textsc{università degli studi di Padova}\\
\textsc{corso di laurea magistrale in Informatica}\\
\textsc{anno accademico 2011 - 2012}\\[3cm]

\begin{tikzpicture}
\foreach \i in {0,10,...,100} {
  \foreach \j in {0,30,...,330} {
    \draw[fill=yellow!\i!red] (\j+.01*\i*\i:.04*\i+.001*\i) circle (.01*\i+.1);
  }
}
\foreach \j in {0,30,...,330} {
  \draw[fill=red!70!green] (\j+1000:4.3) circle (.1cm);
}
\end{tikzpicture}

\end{center}
\end{titlepage}}

\newcommand{\upperborder}{
\begin{tikzpicture}
\clip (0,0) rectangle (24.5, 4);
\foreach \i in {0,1,...,6} {
  \fill[red!40!yellow] (\i*4.5, 0) circle (2.5cm);
  \foreach \j in {0,30,...,180} {
    \fill[red!40!yellow] (\i*4.5,0) +(\j:2.5cm) circle (.7cm);
    \foreach \z in {0,40,...,360} {
      \fill[red!40!yellow] (\i*4.5,0) ++(\j:2.5cm) +(\z:1cm) circle (.2cm);
    }
  }
}

\end{tikzpicture}}

\fancypagestyle{testo}{
\pagenumbering{arabic}
\setcounter{page}{1}
%\lhead{\begin{picture}(0,0)(145,4)\upperborder\end{picture}}
\chead{}
\rhead{}
\lfoot{\footnotesize \textsc{Sistemi Ipermediali}}
\cfoot{\thepage}
\rfoot{}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}
}